\documentclass[11pt]{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage[margin=1in]{geometry}
\usepackage{doi}

\providecommand{\Sym}{\mathrm{Sym}}
\providecommand{\End}{\mathrm{End}}
\providecommand{\Heis}{\mathrm{Heis}}
\providecommand{\Ihat}{\widehat{I}}

\newtheorem{lemma}{Lemma}
\newtheorem{remark}{Remark}

\title{The Oriented Frontier Observable:
a Recursive Q11 Diagnostic for the Angular Generation Split\\[2pt]
\large (internal note, twin to the symmetric-capacity obstruction)}
\author{Cosmochrony programme --- working note}
\date{\today}

\begin{document}
  \maketitle

  \begin{abstract}
    The symmetric-capacity obstruction establishes that the raw shell observable of the angular generation
    split vanishes identically, $\Theta_{\mathrm{raw}}=0$: on a symmetric BFS shell the involution
    $g\mapsto g^{-1}$ pairs opposite central phases and erases the oriented $J_\Pi$-odd component before the
    radial recursion can act.
    This note defines the honest replacement: a frontier transfer observable attached to the directed
    outgoing edges $g\to gs$ of the cascade, whose central increment is the Heisenberg cocycle $a(g)\,s_b$,
    the discrete counterpart of the Baker--Campbell--Hausdorff term $\tfrac{1}{2}[X,Y]$ and hence of the
    dynamical $J_3$~\cite{Beau2026q11,Beau2026q14}.
    The orientation is the recursive arrow $\partial_\tau=\log g$ from the projection origin to increasing
    BFS depth, not an extracted weight.
    The first test is not an amplitude value but the existence of an oriented signal,
    $\langle\Delta A_c\rangle_{\partial^+S_m}\neq 0$, which may fail if a residual automorphism re-pairs the
    frontier; the discriminating anti-bias control is the symmetrised-frontier cancellation
    $\langle\Delta A_c\rangle_{\partial^+S_m\cup\partial^-S_m}=0$.
    No normalisation $\mathcal{N}_A$, no dictionary value, and no $\varepsilon$ appear in this note: it tests
    only whether the recursion produces a non-zero, bias-free oriented signal.
  \end{abstract}

  \textbf{Keywords:} oriented cascade; frontier transfer; Heisenberg cocycle; recursive orientation;
  generation split; anti-bias control; projected capacity.

  \section{Context}

    Q14, \S6~\cite{Beau2026q14} fixes the angular generation split as an amplitude with a factorised
    architecture and defers its absolute value to the physical normalisation of the metaplectic $J_3$
    generator.
    The companion symmetric-capacity obstruction note isolates the two sectors in the O12 fingerprint
    exponent --- the frequency $B_c$ carries the radial capacity $\sigma_c$, the central phase $A_c$ carries
    the angular split --- and shows that a direct evaluation on symmetric BFS shells returns
    $\Theta_{\mathrm{raw}}=0$ identically.
    The obstruction is not the absence of an angular phase; it is that the symmetric shell average is the
    wrong observable for a recursively oriented cascade.

    The present note records the natural exit, prescribed by the oriented cascade arrow
    $\partial_\tau=\log g$~\cite{Beau2026q11} rather than chosen.
    Concretely, the central phase of a single element $g=(a,b,z)$ under the central character $c$ is
    $A_c(g)=c\,z$, so the swept central coordinate $z$ is the oriented symplectic area from the origin to
    $g$.
    Along one outward step $g\to gs$ the Heisenberg law gives $z(gs)=z(g)+z(s)+a(g)\,s_b$, hence
    \[
      \Delta A_c(g,s) = A_c(gs)-A_c(g) = c\big(z(s)+a(g)\,s_b\big),
    \]
    which for the admissible generators ($z(s)=0$) reduces to the pure cocycle $c\,a(g)\,s_b$.
    This is the discrete $\tfrac{1}{2}[X,Y]$ increment: the transfer observable measures exactly the
    dynamical $J_3$ carried by the oriented edge, not a free shear.

  \section{The oriented frontier observable}

    The symmetric-capacity obstruction shows that the raw shell observable cannot
    select an angular branch: on a symmetric BFS shell $S_m$, the involution
    $g\mapsto g^{-1}$ pairs opposite central phases and cancels the oriented
    $J_\Pi$-odd component.
    The Q11-oriented observable is therefore not a shell average.
    It is a frontier transfer observable.

    Fix the identity element $e$ as the projection origin and let $d(e,g)$ be the
    BFS distance in the admissible Cayley graph.
    For each shell $S_m=\{g:d(e,g)=m\}$, define the outgoing and incoming frontiers by
    \[
      \partial^+S_m
      =
      \{(g,s): d(e,g)=m,\ d(e,gs)=m+1\},
    \]
    \[
      \partial^-S_m
      =
      \{(g,s): d(e,g)=m,\ d(e,gs)=m-1\}.
    \]
    No subset of generators is selected.
    The orientation is the recursive arrow from the projection origin to increasing
    BFS depth.

    For a directed frontier edge $g\to gs$, define
    \[
      \Delta A_c(g,s)=A_c(gs)-A_c(g).
    \]
    Since $A_c$ changes sign under the conjugate branch $c\mapsto q-c$,
    $\Delta A_c$ is already $J_\Pi$-odd:
    \[
      \operatorname{Odd}_{J_\Pi}(\Delta A_c)=\Delta A_c.
    \]
    The $J_\Pi$-odd notation is therefore a parity check, not an additional filter.

    The Q11-oriented raw angular profile is
    \[
      \Theta_{\mathrm{Q11}}^{\mathrm{raw}}(m)
      =
      \left\langle
        \frac{2\pi}{q}\Delta A_c(g,s)
      \right\rangle_{\partial^+S_m}.
    \]
    Its cumulative, capacity-normalised form is
    \[
      \Theta_{\mathrm{Q11}}^{\mathrm{oriented}}(n)
      =
      \frac{1}{\Ihat(n)}
      \sum_{m\le n}
      \Theta_{\mathrm{Q11}}^{\mathrm{raw}}(m)\,\Delta\Ihat(m).
    \]
    This quantity is prescribed by the oriented frontier of the cascade, not extracted
    from symmetric capacity data.

    The first test is not the value of the amplitude, but the existence of an oriented
    signal:
    \[
      \Theta_{\mathrm{Q11}}^{\mathrm{raw}}(m)\overset{?}{\neq}0.
    \]
    This test may fail.
    The outgoing frontier is canonical once the origin and admissible generators are
    fixed, but a residual automorphism of the frontier may still re-pair the cocycle
    contributions.

    The required anti-bias control is the symmetrised-frontier cancellation:
    \[
      \left\langle
        \frac{2\pi}{q}\Delta A_c(g,s)
      \right\rangle_{\partial^+S_m\cup\partial^-S_m}
      \overset{!}{=}0.
    \]
    Thus a valid oriented signal must satisfy
    \[
      \left\langle \Delta A_c\right\rangle_{\partial^+S_m}\neq0,
      \qquad
      \left\langle \Delta A_c\right\rangle_{\partial^+S_m\cup\partial^-S_m}=0.
    \]
    Only then can the signal be attributed to the recursive orientation rather than
    to a hidden sampling bias.

  \section{BI parity versus the residual reflection}

    The outgoing-frontier observable isolates a residual obstruction.
    The obstruction is not the BI parity itself.
    It is the reflection
    \[
      \varphi:(a,b,z)\longmapsto(-a,b,-z),
    \]
    which preserves the outgoing frontier and reverses the central cocycle.

    In the O12 fingerprint variables
    \[
      \mathrm{fp}_c(g;x)
      =
      q^{-3/2}
      \exp\left(\frac{2\pi i}{q}(A_c(g)+B_c(g)x)\right),
    \]
    the action of $\varphi$ is
    \[
      \varphi:(A_c,B_c)\longmapsto(-A_c,+B_c).
    \]
    By contrast, the spinorial BI parity of Q14~\cite{Beau2026q14} is the antiunitary structure
    $J_\Pi=HK$.
    In the weight basis Q14 fixes
    \[
      K=\sigma_x\circ\overline{(\cdot)},\qquad H=\sigma_z,\qquad
      J_\Pi^{\mathrm{int}}=HK=\varepsilon\circ\overline{(\cdot)}.
    \]
    Thus the induced action on the fingerprint character is the BI-conjugate action
    \[
      J_\Pi:(A_c,B_c)\longmapsto(-A_c,-B_c).
    \]
    Consequently
    \[
      J_\Pi\neq\varphi.
    \]
    The residual reflection differs from BI parity by the frequency reflection
    \[
      R_b:(A_c,B_c)\longmapsto(A_c,-B_c),
    \]
    namely
    \[
      \varphi=R_b\circ J_\Pi.
    \]

    This distinction is important.
    The residual obstruction is not removed by the statement that the chiral selector is
    $J_\Pi$-odd.
    Since $\varphi$ differs from $J_\Pi$ by $R_b$, the relevant condition is instead
    whether chirality is blind to the frequency sector:
    \[
      [\gamma_5,R_b]=0.
    \]
    Under this condition, $J_\Pi$-oddness implies $\varphi$-oddness.
    Indeed, using $J_\Pi\gamma_5=-\gamma_5J_\Pi$ and assuming
    $[\gamma_5,R_b]=0$, one obtains
    \[
      \varphi\gamma_5
      =
      R_bJ_\Pi\gamma_5
      =
      -R_b\gamma_5J_\Pi
      =
      -\gamma_5R_bJ_\Pi
      =
      -\gamma_5\varphi.
    \]
    Thus $\gamma_5$ is $\varphi$-odd.

    The frontier cocycle $\Delta A_c$ is itself $\varphi$-odd and arrow-odd.
    Therefore, if $[\gamma_5,R_b]=0$, the chiral integrand
    \[
      \gamma_5\,\Delta A_c
    \]
    is $\varphi$-even and arrow-odd.
    It can survive the residual reflection while retaining the required orientation under
    cascade reversal.

    This is only a conditional lifting statement.
    If $[\gamma_5,R_b]\neq0$, or if $R_b$ is not a legitimate frequency-sector
    operator on the chiral carrier, then the $V-A$ selector of Q14 does not lift the
    $\varphi$-obstruction.
    In that case the angular amplitude remains open even after the Q11 frontier
    orientation is imposed.

  \section{Status and protocol}

    This note defines the object and its falsification protocol only.
    The twin script \texttt{q11\_oriented\_frontier.py} returns the three profiles
    $\Theta^{\mathrm{raw}}_{\partial^+}(m)$, $\Theta^{\mathrm{raw}}_{\partial^-}(m)$,
    $\Theta^{\mathrm{raw}}_{\partial^+\cup\partial^-}(m)$, the cumulative capacity-normalised
    $\Theta_{\mathrm{Q11}}^{\mathrm{oriented}}(n)$, and the three status flags
    (\texttt{sign\_reversal\_check}, \texttt{sym\_frontier\_zero\_check},
    \texttt{automorphism\_residual\_status}).
    The sign-reversal check $\Theta^{\mathrm{raw}}_{\partial^-}=-\Theta^{\mathrm{raw}}_{\partial^+}$ is nearly
    tautological once $\partial^-$ is the reversed frontier; it is a consistency check, not the discriminant.
    The discriminant is the symmetrised cancellation: a signal is admissible only if it is non-zero on
    $\partial^+$ \emph{and} zero on $\partial^+\cup\partial^-$.

    The status of the observable is explicit: the orientation is prescribed by $\partial_\tau=\log g$, not
    extracted from symmetric capacity data.
    Neither the normalisation $\mathcal{N}_A$ nor any comparison to a dictionary value is performed here;
    these remain deferred, and become meaningful only if both tests pass.

    The remaining problem is therefore not an amplitude computation.
    It is the operator question whether the chiral grading $\gamma_5$ commutes with the
    frequency reflection $R_b$.
    Only after this condition is established does it make sense to compute a chiral
    frontier signal and to ask for the normalisation $\mathcal{N}_A$.


    \bibliographystyle{plain}
    \bibliography{cosmochrony-bibliography}

\end{document}
